---
title: "MATH 216 Homework 3"
author: "Katherine Hobbs"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(lubridate)
library(Quandl)
library(tidyr)
library(knitr)
library(forcats)
```



## Admistrative:

Please indicate

* Who you collaborated with: Amanda Hotvedt, Kyra Gray
* Roughly how much time you spent on this HW so far: 5
* The URL of the RPubs published URL [here](http://rpubs.com/khobbs/223599).
* What gave you the most trouble: I had trouble understanding the work flow and specific code required for the regressions, but I think I have a better grasp on in now. 
* Any comments you have:

I still need to figure out how to make the testing and training data sets disjoint and whether I need to set a seed for those models (I have noticed that I get a different proportion for correct predictions each time I run the model). 


## Question 1:

We will use a logistic regression model to predict sex. Our metric to rate how well our
model performs will be:

$$
\frac{1}{n}\sum_{i=1}^{n}I(y_i = \widehat{y}_i)
$$

where $I(A)$ is the *indicator function* that is equal to 1 if condition $A$
holds, 0 otherwise. So

* Say user $i$ **is** female, then $y_i=1$
* Say we **predict** user $i$ is female, then $\widehat{y}_i=1$
* In this case $I(y_i =\widehat{y}_i)=1$. 

So what the above formula is reporting is the proportion of users' sex we
correctly predicted.

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
# To keep this exercise simple, let's remove the only 3 users (0.005% of users)
# who did not list a height, define the outcome variable, and add an ID variable
# to distinguish the users
profiles <- read_csv(file="profiles.csv") %>% 
  filter(!is.na(height)) %>% 
  mutate(is_female=ifelse(sex=='f', 1, 0)) %>% 
  tibble::rownames_to_column(var="id")
```



#### a)

Define:

* A *training* set `training` of 2997 users (5% of users). We will train the 
logistic regression model to predict gender using this data. Since we want to 
train the model to tell who is female and who is not, we use the outcome
variable `is_female`.
* A *test* set `test` of the remaining 56,946 users (95% of users). We will test
how good our trained model is using this data. So at first, we will pretend we
don't know the outcome variable `is_female`. We use the above model to make a
prediction of sex for all 56,946 test users, then we use the `is_female` outcome
to rate how well we performed.
* Be sure to incorporate all the insight your garnered in your EDA in HW-2.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
set.seed(76)

profiles <- profiles %>% 
    mutate(religion= ifelse(is.na(religion), "Not Reported", religion))

profiles <- profiles %>% 
  mutate(`Grouped Religion` = fct_recode(religion,
                  "agnosticism" = "agnosticism",
                  "agnosticism" = "agnosticism and laughing about it",
                  "agnosticism" = "agnosticism and somewhat serious about it",
                  "agnosticism" = "agnosticism and very serious about it",
                  "agnosticism" = "agnosticism but not too serious about it",
                  "atheism" = "atheism",
                  "atheism" = "atheism and laughing about it",
                  "atheism" = "atheism and somewhat serious about it",
                  "atheism" = "atheism and very serious about it",
                  "atheism" = "atheism but not too serious about it",
                  "buddhism" = "buddhism",
                  "buddhism" = "buddhism and laughing about it",
                  "buddhism" = "buddhism and somewhat serious about it",
                  "buddhism" = "buddhism and very serious about it",
                  "buddhism" = "buddhism but not too serious about it",
                  "catholism" = "catholicism",
                  "catholism" = "catholicism and laughing about it",
                  "catholism" = "catholicism and somewhat serious about it",
                  "catholism" = "catholicism and very serious about it",
                  "catholism" = "catholicism but not too serious about it",
                  "christianity" = "christianity",
                  "christianity" = "christianity and laughing about it",
                  "christianity" ="christianity and somewhat serious about it",
                  "christianity" = "christianity and very serious about it",
                  "christianity" = "christianity but not too serious about it",
                  "hinduism" = "hinduism",
                  "hinduism" = "hinduism and laughing about it",
                  "hinduism" = "hinduism and somewhat serious about it",
                  "hinduism" = "hinduism and very serious about it",
                  "hinduism" = "hinduism but not too serious about it",
                  "islam" = "islam",
                  "islam" = "islam and laughing about it",
                  "islam" = "islam and somewhat serious about it",
                  "islam" = "islam and very serious about it",
                  "islam" = "islam but not too serious about it",
                  "judaism" = "judaism",
                  "judaism" = "judaism and laughing about it",
                  "judaism" = "judaism and somewhat serious about it",
                  "judaism" = "judaism and very serious about it",
                  "judaism" = "judaism but not too serious about it",
                  "other" = "other",
                  "other" = "other and laughing about it",
                  "other" = "other and somewhat serious about it",
                  "other" = "other and very serious about it",
                  "other" = "other but not too serious about it",
                  "Not Reported" = "Not Reported"))


profiles <- profiles %>% 
  mutate(income_levels =
    ifelse(income %in% -2:0, "Not Reported",
    ifelse(income %in% 0:20000, "Low Income",
    ifelse(income %in% 20001:70000, "Middle Income",
    ifelse(income %in% 70000:1000000, "High Income", " ")))))
  
profiles <- profiles %>%
  mutate(job =ifelse(is.na(job),"Not Reported", job))

training <- profiles[sample(nrow(profiles), 2997), ]

test <- anti_join(profiles, training, by="id")

#I do not yet know how to make these disjoint data sets! 
```



#### b)

Train the logistic regression model to predict sex. i.e. fit a logistic
regression model to the `training` data. Assign this model to an R object called
`predict_sex_model`, then rate how well the model performs on the `training` data.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}

predict_sex_model <- glm(is_female ~`Grouped Religion` + income_levels + job, data = training, family = "binomial")

broom::tidy(predict_sex_model)

predictions <- training %>% 
  select(income_levels, job, `Grouped Religion`, is_female) %>% 
  mutate(p_hat = fitted(predict_sex_model))

predictions <- predictions %>% 
  mutate(correct = ifelse(p_hat >.5 & is_female==1, 1,
                          ifelse(p_hat<0.5 &is_female==0,1,0)))
                          
predictions %>% 
  summarise(Training_prop_correct=mean(correct)) %>% 
  knitr::kable(digits = 4)
```



#### c)

Take `predict_sex_model` and apply it to the `test` data and make a prediction 
for each users' sex, then rate how well the model performs on the `test` data.

**Hint**: What do you think `predict(predict_sex_model, newdata=test,
type="response")` does? The help file is located in `?predict.glm`

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}

test_predictions <- test %>% 
  select(income_levels, job, `Grouped Religion`, is_female) %>% 
  mutate(p_hat = predict(predict_sex_model, newdata=test, type="response"))

test_predictions <- test_predictions %>% 
  mutate(correct = ifelse(p_hat >.5 & is_female==1, 1,
                          ifelse(p_hat<0.5 &is_female==0,1,0)))
                          
test_predictions  %>% 
  summarise(Test_Prop_correct=mean(correct)) %>% 
  knitr::kable(digits = 4)


```



#### d)

Did the model perform better on the `training` data or the `test` data? Why
do you think that is?

The model performed better on the test data, but only marginally! The results were almost the same. 63.73% of predictions were correct for the training data while 63.67% of predictions were correct for the testing data. I would expect the model to perform better on the training data because that was the data used to build the model, while the test data is a new dataset to which the model is being applied. 


## Question 2:

We want to compare the volatility of 

* [Bitcoin](https://www.quandl.com/data/BAVERAGE/USD) prices
* [Gold](https://www.quandl.com/data/BUNDESBANK/BBK01_WT5511) prices

Let our measure of volatility be the relative change from day-to-day in price. 
Let the reference currency be US dollars. Analyze these results and provide
insight to a foreign currency exchanger.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}

bitcoin <- Quandl("BAVERAGE/USD") %>% 
  tbl_df()

gold <- Quandl("BUNDESBANK/BBK01_WT5511") %>% 
  tbl_df()

bitcoin <- bitcoin %>% 
  rename(
    Avg = `24h Average`,
    Total_Volume = `Total Volume`)

bitcoin <- bitcoin %>% 
  mutate(Price_lag=lag(Avg)) %>% 
  mutate(Daily_Price_Change=(Avg - Price_lag)) %>% 
  mutate(Daily_Rel_Price_Change=(Daily_Price_Change/Price_lag))          

gold_recent <- gold %>% 
  filter(Date %within% interval(ymd("2010-01-01"), ymd("2016-12-31")))

gold_recent <- gold_recent %>% 
    mutate(Value_lag=lag(Value)) %>% 
    mutate(Daily_Value_Change=(Value - Value_lag)) %>% 
    mutate(Daily_Rel_Value_Change=(Daily_Value_Change/Value_lag))     

p <- ggplot() +
  geom_line(data=bitcoin, aes(x=Date, y=Daily_Rel_Price_Change, color="Bitcoin")) +
  geom_line(data=gold_recent, aes(x=Date, y=Daily_Rel_Value_Change, color="Gold")) +
  labs(x='Date', y='USD') +
  labs(title="Daily Relative Change of Bitcoin and Gold Values 2010-2016")

p

```





## Question 3:

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
jukebox <- read_csv(file="reed_jukebox.csv")

# Clean certain artists' names:
sigur_ros <- "Sigur Ro\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8c\x93\xa0\xbcs"
bjork <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk"
blue_oyster_cult <- "Blue O\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcyster Cult"
husker_do <- "Hu\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcsker Du\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbc"
bjork_brodsky <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk & Brodsky Quartet"
slagsmalsklubben <- "Slagsma\xfc\xbe\x99\x86\x94\xbc_lsklubben "
bjork_sugarcubes <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk (Sugarcubes)"

jukebox <- jukebox %>%
  mutate(
    artist = ifelse(artist == sigur_ros, "Sigor Ros", artist),
    artist = ifelse(artist == bjork, "Bjork", artist),
    artist = ifelse(artist == blue_oyster_cult, "Blue Oyster Cult", artist),
    artist = ifelse(artist == husker_do, "Husker Do", artist),
    artist = ifelse(artist == bjork_brodsky, "Bjork & Brodsky Quartet", artist),
    artist = ifelse(artist == slagsmalsklubben, "Slagsmalsklubben", artist),
    artist = ifelse(artist == bjork_sugarcubes, "Bjork (Sugarcubes)", artist)
  )
```


Using the Reed College jukebox data, what are the top 10 artists played during
the "graveyard shift" during the academic year? Define

* the "graveyard shift" as midnight to 8am
* the academic year as September through May (inclusive)

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}

jukebox <- jukebox %>% 
  mutate(Date = parse_date_time(date_time, "a b d HMS Y")) %>% 
  mutate(Month=month(Date)) %>% 
  mutate(hour=hour(Date))

graveyard <- jukebox %>%   
  filter(Month<=5 | Month>=9) %>% 
  filter(hour>=24 | hour <=8)
  
graveyard_artist <- graveyard %>% 
  group_by(artist) %>% 
  tally() %>% 
  ungroup() %>% 
  arrange(desc(n))

graveyard_top10 <- graveyard_artist %>% 
  top_n(10)

graveyard_top10 %>% 
  kable()

```





